---
title: "Creating Functions in R Challenge"
author: "Geoffery Mullings"
date: "March 10, 2016"
output: word_document
---

```{r}
mymean <- function(Y)
  
{
  
  Ybar <- sum(Y) / length(Y)
  
  return(Ybar)
  
}

x = c(1,3,10)
mymean(x)
```

Not surprisingly, the result is the same as the one obtained using the mean function. More generally, a function can take several arguments, but it has to return only one outcome, which could be a list of items. The function we defined above is quite simple and it has several limitations: 1) it does not take into account that the series might have NAs...

```{r}
x = c(1, 3, 10, NA)
mymean(x)
```

...and 2) it does not calculate the mean of each column in case there are several. As an exercise, modify the mymean function to accomodate for these issues.

```{r}
mymean <- function(Y, na.rm = TRUE)
     
 {
  N=ncol(Y)
     if (na.rm==TRUE) {Y=na.exclude(Y)} #Remove NAs from the variable
     
     if (!is.null(N) && length(N) > 1) {for (k in 1:N) #Makes sure the number of columns isn't none.
       totals <- sum(Y[[k]])/length(Y[[k]])
     
     Ybar <- sum(totals) / length(totals) 
     
     return(Ybar)}
     
     else
     
     {Ybar <- sum(Y) / length(Y)
     
     return(Ybar)}
     
}

mymean(x, na.rm=FALSE)

mymean(x)

y = c(1,3,10)

mymean(y)
mean(y)

Data = as.data.frame(cbind(c(1, 3, 10, 50), c(4, 7, 10, 1), c(9, 2, 9, 4)))
mymean(Data)
mean(Data)

```

G.M.