---
title: "Writing A Standard Deviation Function In R"
author: "Geoffery Mullings"
date: "March 16, 2016"
output: word_document
---

Creating functions in R can be a convenient way to automate repetitive processes in analysis, improiving efficiency.

Here's a simple replication of the base statistics package's standard deviation function. First, an example of the base function in action:

```{r}
test = c(1, 5, 64, 22)
sd(test)
```

Now let's replicate the function, in the process giving step by step instructions on how to calculate a standard deviation.

```{r}
mySD = function(x){
themean = mean(x) #Calculate the mean of the observations
N = length(x) #N is a variable to make accessing the count of observations simple.
distmean = {} #Initializing the variable that will store our distances from the mean.
for (i in 1:N) {distmean = c(distmean, x[i]-themean)} #Looping over the observations, these commands pull the distance of each observation from the mean and stores the value in our "distmean" variable.
underroot = sum(distmean^2)/(N-1) #Underroot is the variable that will be square rooted, it contains the sum of our squared distances over the number of observations minus 1 - to account for the limited degrees of freedom.
return(sqrt(underroot))}
```

Does it work?

```{r}
sd(test)
mySD(test)
```

Doors are opened when you learn how to write functions in R.
